---
title: "ST558 Project 1"
author: "Noel Hilliard"
date: "6/5/2020"
output:
  rmarkdown::github_document
  toc:true
---

Project Objective: This project involves creating a vignette (an HTML file with a table of contents) about reading in data in JSON format and exploring it (summaries and graphs - more details on this later). You will then create a
blog post linking to your vignette.All project work should be done in a github repo. 


# Functions

```{r Install Packages, echo = TRUE, message = FALSE}

#install.packages("httr")

library("httr")

#install.packages("jsonlite")

library("jsonlite")
```

## Franchise Data from API
```{r franchise function, echo = TRUE}
getFranchise <- function(){

  base_url <- "https://records.nhl.com/site/api"
  tab_name <- "franchise"
  full_url <- paste0(base_url, "/", tab_name)

  get_franchise <- GET(full_url)

  get_franchise_text <- content(get_franchise, "text")

  get_franchise_json <- fromJSON(get_franchise_text, flatten = TRUE)

  return(data.frame(get_franchise_json))
}

```


## Franchise Team Totals Data from API
```{r FranchiseTeamTotals, echo = TRUE}
getTeamTotals <- function(){
  
  base_url <- "https://records.nhl.com/site/api"
  tab_name <- "franchise-team-totals"
  full_url <- paste0(base_url, "/", tab_name)

  get_TeamTotals <- GET(full_url)

  get_TeamTotals_text <- content(get_TeamTotals, "text")

  get_TeamTotals_json <- fromJSON(get_TeamTotals_text, flatten = TRUE)
  
  return(data.frame(get_TeamTotals_json[[1]]))
  
}
```

## Franchise Season Records Data from API
```{r FranchiseSeasonRecords, echo = TRUE}
getSeasonRecords <- function(FranchiseID){
  
 
  base_url <- "https://records.nhl.com/site/api"
  tab_name <- "franchise-season-records?cayenneExp=franchiseId="
  full_url <- paste0(base_url, "/", tab_name, FranchiseID)

  get_SeasonRecords <- GET(full_url)

  get_SeasonRecords_text <- content(get_SeasonRecords, "text")

  get_SeasonRecords_json <- fromJSON(get_SeasonRecords_text, flatten = TRUE)
 
  return(data.frame(get_SeasonRecords_json))
  
}

```

## Franchise Goalie Records Data from API

```{r GoalieRecords, echo = TRUE}
getGoalieRecords <- function(FranchiseID){
  
  base_url <- "https://records.nhl.com/site/api"
  tab_name <- "franchise-goalie-records?cayenneExp=franchiseId="
  full_url <- paste0(base_url, "/", tab_name, FranchiseID)

  get_GoalieRecords <- GET(full_url)

  get_GoalieRecords_text <- content(get_GoalieRecords, "text")

  get_GoalieRecords_json <- fromJSON(get_GoalieRecords_text, flatten = TRUE) 
  
  return(data.frame(get_GoalieRecords_json))
}

```

## Franchise Sakter Records Data from API

```{r SkaterRecords, echo = TRUE}
getSkaterRecords <- function(FranchiseID){
  
  base_url <- "https://records.nhl.com/site/api"
  tab_name <- "franchise-skater-records?cayenneExp=franchiseId="
  full_url <- paste0(base_url, "/", tab_name, FranchiseID)

  get_SkaterRecords <- GET(full_url)

  get_SkaterRecords_text <- content(get_SkaterRecords, "text")

  get_SkaterRecords_json <- fromJSON(get_SkaterRecords_text, flatten = TRUE) 
  
  return(data.frame(get_SkaterRecords_json))
}

```



