---
title: "ST558 Project 1"
author: "Noel Hilliard"
date: "6/5/2020"
output:
  rmarkdown::github_document:
    toc: true
    
---

Project Objective: This project involves creating a vignette (an HTML file with a table of contents) about reading in data in JSON format and exploring it (summaries and graphs - more details on this later). You will then create a blog post linking to your vignette.All project work should be done in a github repo. 

# JSON Data

## What is it?

JavaScript Object Notation (JSON) data is a lightweight data-interchange format used to store data. According to [JSON.org](https://www.json.org/json-en.html), JSON data has two data structures that are recognized by all programming languages: (1) A collection of name/value pairs (2) An ordered list of values. The first data structure is recognized as "an object, record, struct, dictionary, hash table, keyed list, or associative array." The second data structure is recognized as "an array, vector list, or sequence." There are seven value types: string, number, object, array, true, false, and null (https://javaee.github.io/tutorial/jsonp001.html) JSON data is a text format that does not depend on language so it is compatible with several languages including C, C++, C#, Java, JavaScript, Perl, and Python.[(JSON.org)](https://www.json.org/json-en.html)

## Where does it get used?

JSON data is used for a variety of applications. JSON data format can be used for browser extensions and websites that are written in JavaScript. It can be used as an alternate data format for XML to transmit data between the client and server in web applications. Other web applications include the use of JSON data in programming of Web Services and APIs to provide access to public data. Lastly, JSON data is used for serialization and transmission of the structured data over the network.(https://blog.eduonix.com/web-programming-tutorials/learn-use-json/)


## Why is it a good way to store data?

JSON data is a good way to store data because it is a text only format. The format can be conveniently transferred with servers and JSON is a data format that can be used by any programming language. (https://www.w3schools.com/js/js_json_intro.asp)

# JSON Packages in R

The three major packages to read JSON data into R are `rJSON`, `RJSONIO`, and `RJSONlite`. All three packages have two identical functions to convert between JSON data and `R` objects: `toJSON` and `fromJSON`. In addtion to these functions, `rJSON` has a function called `newJSONParser` to convert JSON objects into R objects. `RJSONIO` has an additional function called `asJSVars` to serialize R objects. The last package, `RJSONlite` is the one I chose to use for this vignette because it can interact with APIs, building pipelines, and streaming data. (http://anotherpeak.org/blog/tech/2016/03/10/understand_json_3.html)


# Functions

To get started, load the `httr` and `jsonlite` libraries. The `httr` library will be used to contact the NHL records 'Franchise' API. The `JSONlite` library will be used to convert JSON data to `R` objects. 

```{r Install Packages, echo = TRUE, message = FALSE}

#install.packages("httr")

library("httr")

#install.packages("jsonlite")

library("jsonlite")
```

There are five functions in this vignette:

* `getFranchise`
* `getTeamTotals`
* `getSeasonRecords`
* `getGoalieRecords`
* `getSkaterRecords`

Each function uses a `GET` request to contact the host [NHL records 'Franchise' API]() by initializing the variables for the API call. Each function returns a data frame of results for a specific table. Next, the data is parsed through to create a useful `R` object by changing the raw data from the API call into JSON format using the `httr` function called `content`. Finally, the parsed JSON data is converted to a dataframe and returned by the function.

## Franchise Data from API

The `getFranchise` function returns `id`, `firstSeasonId`, and `lastSeasonId` and name of every team in the history of the NHL from the API for each franchise in a dataframe.

```{r franchise function, echo = TRUE}
getFranchise <- function(){

  base_url <- "https://records.nhl.com/site/api"
  tab_name <- "franchise"
  full_url <- paste0(base_url, "/", tab_name)

  get_franchise <- GET(full_url, timeout = 1000)

  get_franchise_text <- content(get_franchise, "text")

  get_franchise_json <- fromJSON(get_franchise_text, flatten = TRUE)

  return(data.frame(get_franchise_json))
}

```

The `getFranchise` function returns a dataframe with the following columns:

```{r getFranchise dataframe, echo = TRUE, message = FALSE}
names(getFranchise())
```

## Franchise Team Totals Data from API

The `getTeamTotals` function returns total stats for every franchise from the API.

```{r FranchiseTeamTotals, echo = TRUE, message = FALSE}
getTeamTotals <- function(){
  
  base_url <- "https://records.nhl.com/site/api"
  tab_name <- "franchise-team-totals"
  full_url <- paste0(base_url, "/", tab_name)

  get_TeamTotals <- GET(full_url, timeout = 1000)

  get_TeamTotals_text <- content(get_TeamTotals, "text")

  get_TeamTotals_json <- fromJSON(get_TeamTotals_text, flatten = TRUE)
  
  return(data.frame(get_TeamTotals_json))
  
}
```

The `getTeamTotals` function returns a dataframe with the following columns:

```{r getTeamTotals, echo = TRUE, message = FALSE}
names(getTeamTotals())
```


## Franchise Season Records Data from API

The `getSeasonRecords` function returns season records for the specified `FranchiseID`.

```{r FranchiseSeasonRecords, echo = TRUE}
getSeasonRecords <- function(FranchiseID){
  
 
  base_url <- "https://records.nhl.com/site/api"
  tab_name <- "franchise-season-records?cayenneExp=franchiseId="
  full_url <- paste0(base_url, "/", tab_name, FranchiseID)

  get_SeasonRecords <- GET(full_url,timeout = 1000)

  get_SeasonRecords_text <- content(get_SeasonRecords, "text")

  get_SeasonRecords_json <- fromJSON(get_SeasonRecords_text, flatten = TRUE)
 
  return(data.frame(get_SeasonRecords_json))
  
}

```

Here, the `getSeasonRecords` function returns a dataframe for the specified franchise with the following columns. Here is an example of the columns when 2 is specified as the `FranchiseID`:

```{r getSeasonRecords dataframe, echo = TRUE, message = FALSE}
names(getSeasonRecords(2))
```

## Franchise Goalie Records Data from API

The `getGoalieRecords` function returns goalie records for the specified `FranchiseID`.

```{r GoalieRecords, echo = TRUE}
getGoalieRecords <- function(FranchiseID){
  
  base_url <- "https://records.nhl.com/site/api"
  tab_name <- "franchise-goalie-records?cayenneExp=franchiseId="
  full_url <- paste0(base_url, "/", tab_name, FranchiseID)

  get_GoalieRecords <- GET(full_url, timeout = 1000)

  get_GoalieRecords_text <- content(get_GoalieRecords, "text")

  get_GoalieRecords_json <- fromJSON(get_GoalieRecords_text, flatten = TRUE) 
  
  return(data.frame(get_GoalieRecords_json))
}

```

Here, the `getGoalieRecords` function returns a dataframe for the specified franchise with the following columns. Here is an example of the columns when 2 is specified as the `FranchiseID`:

```{r getGoalieRecords, echo = TRUE, message = FALSE}
names(getGoalieRecords(2))
```

## Franchise Skater Records Data from API

The `getSkaterRecords` function returns skater records for the specified `FranchiseID`.

```{r SkaterRecords, echo = TRUE}
getSkaterRecords <- function(FranchiseID){
  
  base_url <- "https://records.nhl.com/site/api"
  tab_name <- "franchise-skater-records?cayenneExp=franchiseId="
  full_url <- paste0(base_url, "/", tab_name, FranchiseID)

  get_SkaterRecords <- GET(full_url, timeout = 1000)

  get_SkaterRecords_text <- content(get_SkaterRecords, "text")

  get_SkaterRecords_json <- fromJSON(get_SkaterRecords_text, flatten = TRUE) 
  
  return(data.frame(get_SkaterRecords_json))
}

```

Here, the `getSkaterRecords` function returns a dataframe for the specified franchise with the following columns. Here is an example of the columns when 2 is specified as the `FranchiseID`:

```{r getSkaterRecords, echo = TRUE, message = FALSE}
names(getSkaterRecords(2))
```

# Contingency Tables 

This section covers using the functions above to create contingency tables of the NHL data. Contingency tables are tables used to show the frequency of categorical variables.

## Active and Inactive Players in Goalie Records for Select Franchises

Below, the goalie records of select franchises are binded together into 1 dataframe and the 2 categorical variables of interest are: `ActivePlayer` and `franchiseName`. Based on the contingincy table, all of the selected franchises have more inactive players than active players.

```{r goalie records counts table, echo = TRUE, message = FALSE}
library(knitr)
#Create dataframe of goalie records for select franchises
GoalieRecords <- rbind(getGoalieRecords(5),getGoalieRecords(10), getGoalieRecords(20), getGoalieRecords(25), getGoalieRecords(15), getGoalieRecords(22))

#factor columns of interest
GoalieRecords$data.activePlayer <- factor(GoalieRecords$data.activePlayer)
GoalieRecords$data.franchiseName <- factor(GoalieRecords$data.franchiseName)

#create table
kable(table(GoalieRecords$data.franchiseName,GoalieRecords$data.activePlayer), 
      col.names = c("Inactive Player", "Active Player"))
```

## Positions in Skater Records for Select Franchises

The dataframe for this contingency table was created in a similar way to the table above by binding together the skater records of select franchises and factor

```{r position by franchise, echo TRUE, message = FALSE}

#Create dataframe of goalie records for select franchises
SkaterRecords <- rbind(getSkaterRecords(5), getSkaterRecords(10), getSkaterRecords(20),getSkaterRecords(25), getSkaterRecords(15), getSkaterRecords(22))

#factor columns of interest
SkaterRecords$data.franchiseName <- factor(SkaterRecords$data.franchiseName)
SkaterRecords$data.positionCode <- factor(SkaterRecords$data.positionCode)

#create table
kable(table(SkaterRecords$data.franchiseName, SkaterRecords$data.positionCode))

```

# Numeric Summaries

## Numeric Summary of Select Columns from Skater Records

```{r stat sum skater records function, echo = TRUE, message = FALSE}
library(dplyr)

SkatersTable <- function(position){
  data <- getSkaterRecords(5) %>% filter(data.positionCode == position) %>% select(data.assists,data.gamesPlayed, data.goals, data.points, data.seasons)
  kable(apply(data, 2, summary))
}

```

```{r position numeric summary L, echo = TRUE, message = FALSE}
SkatersTable("L")
```

```{r position numeric summary C, echo = TRUE, message = FALSE}
SkatersTable("C")
```

```{r position numeric summary D, echo = TRUE, message = FALSE}
SkatersTable("D")
```

```{r position numeric summary R, echo = TRUE, message = FALSE}
SkatersTable("R")
```

# Plots

## Barplot

```{r Skater bar plot, echo = TRUE, message = FALSE}
library(ggplot2) #load ggplot for plots

#dataframe of subset of skater records 
SkaterRecords <- rbind(getSkaterRecords(5), getSkaterRecords(10), getSkaterRecords(20),getSkaterRecords(25), getSkaterRecords(15), getSkaterRecords(22))

#rename factor levels
SkaterRecords$data.activePlayer <- factor(SkaterRecords$data.activePlayer)
player_new <- c("Inactive Player", "Active Player")
levels(SkaterRecords$data.activePlayer) <- player_new

#
g <- ggplot(data = SkaterRecords, aes(x = data.franchiseName))
g + geom_bar(aes(fill = data.activePlayer), position = "dodge") + labs(x = "Franchise Name", fill = "") + theme(axis.text.x = element_text(size = 10, angle = 90))
```

## Boxplot

```{r losses boxplot, echo = TRUE, message = FALSE}

#create dataframe of select franchises from goalie records
GoalieRecords <- rbind(getGoalieRecords(5),getGoalieRecords(10), getGoalieRecords(20), getGoalieRecords(25), getGoalieRecords(15), getGoalieRecords(22))

#create barplot
g <- ggplot(GoalieRecords, aes(x = data.franchiseName, y = data.losses))
g + geom_boxplot() + theme(axis.text.x = element_text(size = 10, angle = 90)) + labs(x = "Franchise Name", y = "Losses")


```

## Scatter Plot

```{r scatter plot, echo = TRUE, message = FALSE}

SkaterRecords <- rbind(getSkaterRecords(23), getSkaterRecords(16), getSkaterRecords(3), getSkaterRecords(6))

g <- ggplot(SkaterRecords, aes(x = data.gamesPlayed, y = data.mostGoalsOneSeason))
g + geom_point(aes(color = data.franchiseName)) + labs(x = "Games Played", y = "Most Goals in One Season", color = "Franchise Name")

```

